A Bayesian Approach to Knowledge Graph Node Embeddings

### 2017-08-16 fractal-plr/src/fractal.prl-final.tex
...

\noindent The word \emph{fractal} is not precisely defined\footnote{Mandelbrot originally defined fractals using Haussdorf dimension, but later stated that he preferred the word to be not precisely defined.}, but typically refers to objects with self-similarity, a non-integer Hausdorff dimension, and an infinitely fine structure.

>>>>

\noindent The word \emph{fractal} is not precisely defined \footnote{Mandelbrot originally defined fractals using Haussdorf dimension, but later stated that he preferred the word to be not precisely defined.}, but typically refers to objects with self-similarity, a non-integer Hausdorff dimension, and an infinitely fine structure.

### 2017-08-16 kgmotifs/kgmotifs.tex
...

@@

@@

@@

\begin{description}
 \item[the graph dimensions] We first store $|V_G|$, $|E_G|$ and $|R_G|$ using the generic code $L^{\N}(\cdot)$. We also store the size of $\I$.
 \item[the pattern] Let $M'$ be the unlabeled graph representation of $M$: that is, $M'$ is a plain graph (not a pattern), $V_{M'}$ contains only the first $V_{M'}$ natural numbers and similarly for $R_{M'}$. Otherwise, $M$ is isomorphic to $M'$: the node and edge labels of $M$ can be mapped to those of $M'$ so that $M$ only contains an edge $E$ if and only if $M'$ contains the edge resulting from mapping $E$'s symbols. 
 \item[the template] This is the graph, minus all links occurring in  instances. Let $E_G'$ be $E_G$ minus any link occurring in any member of $\cal I$. We then store $(V_G, E_G')$ using $L^\text{base}(\cdot)$.
 \item[the instances] In order to reconstruct the graph from the template, we need to know which nodes correspond to instances of the motif, and in what order. For each node in $M'$, we create a list of length $\cal |I|$ of the nodes in $G$ to which it maps for each instance. For the non-variable nodes of the pattern, this will be a list simply repeating one node node of $G$. For the variable nodes, it will be a list containing some subset of the nodes in $G$.
 We store each list using the DM code. It may seem wasteful to store a list repeating a single element (instead of storing a labeled version of the pattern), but under the DM code such lists will be stored efficiently, using only a small, constant number of bits. The more varied the list, the more bits will be required to store it.
 \item[the edge labels] We now have enough information to reconstruct the graph structure. All that remains to fully reconstruct $G$ is to label the links of the motif instances. As for the nodes, we create one list of length $\cal I$ for each type of link in $M'$, which we store using the DM code.
\end{description}

@@

We can define a knowledge graph by three length-$m$ integer sequences: $S$, $P$, $O$, with $\{(S_j, P_j, O_j)\}_j$ the graph's tripleset. If the graph satisfies degree sequence $D$, then we know that $S$ should contain node $j$ $D^\text{out}_j$ times, $P$ should contain relation $r$ $D^\text{rel}_r$ times and $O$ should contain node $j$ $D^\text{in}_j$ times.  Let ${\cal S}_D$ be the set of all such triples of integer sequences satisfying $D$. We have 
\[
|{\cal S}_D| =
 {m \choose {D_1^\text{out}, \ldots, D_n^\text{out}} }
 {m \choose {D_1^\text{rel}, \ldots, D_n^\text{rel}} }
 {m \choose {D_1^\text{in}, \ldots, D_n^\text{in}} } \text{.}
\]

@@

We can thus use 
\[
p^\text{EL}_D(G) =  \frac{m!}{{m \choose {D_1^\text{out}, \ldots, D_n^\text{out}} }
 {m \choose {D_1^\text{rel}, \ldots, D_n^\text{rel}} }
 {m \choose {D_1^\text{in}, \ldots, D_n^\text{in}} }}
\]
as an approximation for the DS model. We call this the edge-list (EL) model. It always lower-bounds the true degree sequence model, since it affords some probability mass to graphs that cannot exist.\footnotemark ~Experiments in the classical motif setting have shown that the ES model is an acceptable proxy \cite{bloem2017large}, especially considering the extra scalability it affords.

\footnotetext{Note that we cannot simply think of $p^\text{EL}$ as a uniform model for graphs containing multiple links, since in we can only divide by $m!$ if we know that all triples are unique.}

@@

In this section, we show some experiments that are not intended to substantiate any claims about our method, but rather to illustrate the variety of uses for succesfull motif detection.

@@

>>>>

\usepackage{easylist}
\usepackage{soul}

@@

\section{TODO}

\paragraph{Code}

\begin{easylist}[itemize]
§ Create fast KGraph implements DTGraph$<$Integer, Integer$>$.
§§ \st{memory-back implementation}
§§ \st{Load RDF data (from HDT)}
§§ \st{Copy all tests from DTGraph, test}
§§ Create Disk-backed version.
§ Implement null model.
§ Implement sampling code. 
§§ $\to$ Implement Nauty for knowledge graphs
§§§ Test \emph{really} thoroughly.
§§ Implement sampling.
§§ Test throughly. 
§ Implement \textbf{motif code}
§§ Tests:
§§§ Implement random sampler. Check that random samples do not have significant motifs.
§§§ Sum probabilities for large sample of small graphs. Check that they do not sum to more than one.
§§§ ???
\end{easylist}

\paragraph{Paper}

\begin{easylist}[itemize]
§ Write preliminaries
§§ Code and MDL
§§ DM code, $\N$ code
§ Write hypothesis test method
§ Experiments
§ Conclusion
\end{easylist}

@@

\paragraph{Common codes}

Extended DM code: test if necessary

@@

\begin{description}
 \item[the graph dimensions] We first store $|V_G|$, $|E_G|$ and $|R_G|$ using the generic code $L^{\N}(\cdot)$. We also store the size of $\I$.
 \item[the pattern] Let $M'$ be the unlabeled graph representation of $M$: that is, $M'$ is a plain graph (not a pattern), $V_{M'}$ contains only the first $V_{M'}$ natural numbers and similarly for $R_{M'}$. Otherwise, $M$ is isomorphic to $M'$: the node and edge labels of $M$ can be mapped to those of $M'$ so that $M$ only contains an edge $E$ if and only if $M'$ contains the edge resulting from mapping $E$'s symbols. 
 \item[the template] This is the graph, minus all links occurring in  instances. Let $E_G'$ be $E_G$ minus any link occurring in any member of $\cal I$. We then store $(V_G, E_G')$ using $L^\text{base}(\cdot)$.
 \item[the instances] In order to reconstruct the graph from the template, we need to know which nodes correspond to instances of the motif, and in what order. For each node in $M'$, we create a list of length $\cal |I|$ of the nodes in $G$ to which it maps for each instance. For the non-variable nodes of the pattern, this will be a list simply repeating one node node of $G$. For the variable nodes, it will be a list containing some subset of the nodes in $G$.
 We store each list using the extended DM code. It may seem wasteful to store a list repeating a single element (instead of storing a labeled version of the pattern), but under the extended DM code such lists will be stored efficiently, using only a small, constant number of bits. The more varied the list, the more bits will be required to store it.
 \item[the edge labels] We now have enough information to reconstruct the graph structure. All that remains to fully reconstruct $G$ is to label the links of the motif instances. As for the nodes, we create one list of length $\cal I$ for each type of link in $M'$, which we store using the DM code.
\end{description}

@@

We can define a knowledge graph by three length-$m$ integer sequences: $S$, $P$, $O$, with $\{(S_j, P_j, O_j)\}_j$ the graph's tripleset. If the graph satisfies degree sequence $D$, then we know that $S$ should contain node $j$ $D^\text{out}_j$ times, $P$ should contain relation $r$ $D^\text{rel}_r$ times and $O$ should contain node $j$ $D^\text{in}_j$ times.  Let ${\cal S}_D$ be the set of all such triples of integer sequences satisfying $D$. We have 
\[
|{\cal S}_D| =
 {m \choose {D_1^\text{out}, \ldots, D_n^\text{out}} }
 {m \choose {D_1^\text{rel}, \ldots, D_{|R_G|}^\text{rel}} }
 {m \choose {D_1^\text{in}, \ldots, D_n^\text{in}} } \text{.}
\]

@@

We can thus use 
\[
p^\text{EL}_D(G) =  \frac{m!}{{m \choose {D_1^\text{out}, \ldots, D_n^\text{out}} }
 {m \choose {D_1^\text{rel}, \ldots, D_{|R_G|}^\text{rel}} }
 {m \choose {D_1^\text{in}, \ldots, D_n^\text{in}} }}
\]
as an approximation for the DS model. We call this the edge-list (EL) model. It always lower-bounds the true degree sequence model, since it affords some probability mass to graphs that cannot exist.\footnotemark ~Experiments in the classical motif setting have shown that the EL model is an acceptable proxy for the DS model \cite{bloem2017large}, especially considering the extra scalability it affords.

\footnotetext{Note that we cannot simply think of $p^\text{EL}$ as a uniform model for graphs containing multiple links, since we can only divide by $m!$ if we know that all triples are unique.}

@@

In this section, we show some experiments that are not intended to substantiate any claims about our method, but rather to illustrate the variety of uses for succesful motif detection.

@@

\appendix

\section{Appendix}

\subsection{The canonical isomorphism algorithm for knowledge graphs}

During sampling, we want to re-order the nodes in a sampled pattern to some canonical ordering. This way, when we sample isomorphic patterns, we recognize it, and group their instances together. The \emph{nauty} algorithm is the standard solution for findings such \emph{canonical isomorphs}. As we are sampling partially labeled, directed graphs, we have a lot of information that can speed up canonicalization. Since this steop is usually the bottleneck in sampling, it's important to perform it as efficiently as possible. Below, we provide a very succinct description of how nauty works, when applied to our patterns.

Let $V'_M$ and $R'_M$ represent the labels of the variable nodes and links in $M$. Pattern $M$ is an isomorphism of pattern $M'$ if there exist bijections $\sigma_v : V'_M \times V'M$ and $\sigma_r: \R'_M \times \R'_M$ such that applying these to the variable elements of $M$ yields $M'$ (under graph equivalence). The set of all patterns isomorphic to $G$ is the isomorphism class $[G]$. A function $C(P)$ is a canonicalization function if it produces for all any pattern $M$ in $[M]$ the same, canonical member of $M$.

Note that a very simple canonicalization algorithm would be to pass the unlabeled graph to the nauty algorithm, and use its canonical ordering on the labeled graph. However, the labels can make the canonicalization procedure significantly faster.

The key principle behind the nauty algorithm is the use of a \emph{colouring}: a partition of the nodes of a graph. A \emph{safe colouring} is defined as a partition $\pi$ of the nodes of the graph such that any permutation that maps nodes from different partitions to one another is guaranteed \emph{not} to be an automorphism. No guarantees are made about nodes in the same cell. For instance, the partition that puts all nodes in the same cell is always safe. If two nodes have different degrees, we can put them in different cells and be sure that we have a safe colouring. Given one safe colouring $\pi$, if node $a$ from partition $i$ has a neighbor in partition $j$ and node $b$ from partition $j$ has no such node, we can \emph{refine} the partition $\pi$ by separating $i$ into a new partition (one contain $a$ and one containing $b$).

The nauty algorithm operates on a graph with a given colouring (with the unit colouring used by default). It first refines the colouring as much as possible, using degree information, and propagating the distinguishing features through the graph. Once the most refined colouring has been achieved, it searches through all autorphisms of the graph to produce a canonical example (using various optimizations to prune the search tree).

In our case, we use the colouring to indicate what our labeling tells us about the colouring of the nodes.
Note the following:
\begin{easylist}
§ Labeled nodes in the pattern can be placed in a discrete cell of the colouring.
§ Variable nodes should be placed into a single cell in the partition.
§ We define a partition over nodes \emph{and} relation labels. 
joh\end{easylist}

### 2017-08-16 kgpsr/kgpsr.tex
...

@@

\newcommand{\R}{{\mathbb R}}
\newcommand{\fw}[1]{\text{\texttt{#1}}}

@@

Let $x_i \in \R^k$ be the embedding chosen for node $i \in V$. Let $f_r$ be a parametrized function from $\cal F$, representing relation $r \in R$. For our method, we will let $\cal F$ be the set of affine functions, each parametrized by a transformation matrix $A_r \in \R^{k\times k}$, and a translation vector $t_r\in \R^k$. Thus our model contains $|V|$ vectors in $\R^k$, and $|R|$ functions each defined by a matrix $A_r$ and vector $t_r$.

To formalize the intuitions expressed in the introduction, we first express the constraints in a graphical model. Let $E^r_\text{subject} = \{s \mid \exists o, (s, r, o) \in E\}$ and $E^{r, o}_\text{object} = \{ s \mid (s, r, o) \in E\}$. Figure~\ref{figure:graphical-model} shows the relations between the different random variables. Interestingly, this model contains \emph{no observed variables}. The data is used purely to inform the topology of the graphical model.\footnotemark The model also gives us the priors: the prior probability for the embeddings, for which we will use the standard normal distribution: $\mu = 0$ and $\Lambda = \bf{1}$ and the prior $W_t$ on the

\footnote{Modeling the data as observations would lead to tensor-factorization approaches like ReSCAL \cite{}.}

% TODO Read up on Bayes and conjugate priors. Get a better sense of what priors I should use.

\[
\argmin_{\{x_i\}, \{f_r\}, \sigma} \prod_{(s,r,o) \in E_G} f_r(N_{x_s,\sigma})(x_o)
\]

If we have $(s, p, o) \in G$, then we have:
\begin{align*}
p(X_o \mid F_p = f_p) = f_p(p(X_s))(X_o) \\
p(X_s \mid F_p = f_p) = f_p^{-1}(p(X_o))(X_s) 
\end{align*}

Where $X_s$ are random variables over $\R^k$ representing the embedding of $s$, and $F_p$ is a random variable over $\cal F$ representing the function chosen for relation $p$. What this states, is that for a given choice of function embedding $f_p$, the probability distribution over $X_o$ can be found by taking the probability distribution over $X_s$, and transforming it by $f_p$. Starting with the joint distribution, and marginalizing out $F_p$:

\begin{align*}
p(X_o) = \int p(F_p = f_p) f_p(p(X_s))(X_o) df_p\\
p(X_s) = \int p(F_p = f_p) f_p(p(X_o))(X_s) df_p\\
\end{align*}

Which is a recursive, cyclic definition. Use KL divergence instead, to measure error.

@@

>>>>

\newcommand{\R}{{\mathbb R}}
\newcommand{\cN}{{\mathcal N}}
\newcommand{\fw}[1]{\text{\texttt{#1}}}
\newcommand{\p}{\text{.}}

@@

\begin{easylist}
§ Strict Bayesian approach doesn't seem to work: no data, no posterior. But, posterior over latent variables. Latent variables are embeddings and tranformations, \emph{observed are the memberships/responsibilities }. So we \emph{do} have observed variables?
§§ Optimize for the posterior over the connections... This would tie in better with the link prediction use case.  
§§ If the structure of the grpah is the latent variable, we should be bale to sample a structure from the model.
§ Take inspiration from the fractal model start with $X_0 ~ N(0, I)$ as embedding for all
nodes and iterate $X_1$ is $X_0$ transformed by all translations, for all points.
§ It doesn't look like the constraints of just a probability model are sufficient to determine the problem. It seems most elegant to assume that the embeddings are (isometric) Gaussians, and to optimize for minimal KL divergence to the transformed ones.

\end{easylist}

@@

Let $T_{spo}$ be a random bolean variable indicating whether the triple $\langle s, p, o\rangle$ is true. Let $T^u$ be the set of all such variables (for all combination of $s$, $p$, $o$) and let $T$ be the set of such variables for the observed triples. We will assume that the observed triples we samples from the set of all true triples by some unspecified distribution $p(G| G^u)$. This graph is fully specified by a set of node-embeddings $x_i$ and relation-maps $f_p$: we have $f_p(x_s) = x_o$ is and only if the triple $s, p, o$ is true. Thus, the embeddings specify the graph. We assume that the observed triples $T$ constitute a sequence of i.i.d. draws from the set of all triples satisfying the embeddings.

To develop our model, we will first assume we know a value $x_i \in \R^k$ for each node $i$ in the knowledge graph, and a function $f_p$ for each predicate $p$. We will choose affine functions as our family, so that we can say (with some abuse of notation) $f_p \in \R^{k\times k} \times \R^k$. For such a fully known model, we will say that a triple is true if and only if $f_p$ is mapped exactly onto $x_s$:
\[
p(T_{spo} = \text{true} \mid f_p, x_s, x_o) = \begin{cases} 1 & \text{if} f_p(x_s) = x_o \\ 0 & \text{otherwise} \end{cases} \p
\]

This model suffers from the same problem as TransE: we cannot represent multiple relations. Interestingly, the fact that we do not observer $x_s$, $x_o$ and $f_p$ directly allows us to solve this problem. We express our uncertainty about these values in probability. If $s$ is mapped to several points by $p$, we are thus \emph{uncertain} about which point $f_p$ should map $x_s$ to. Expressing this uncertainty as probability as well allows as to retrieve a probability distribution over the model parameters that let $x_s$ map to deveral different $x_o$'s under $f_p$.

We start dy removing the assumption that we know the model parameters. We express our uncertainty as probabilities: $p(F_p = f_p)$, $p(X_s = x_s)$, $p(X_o = x_o)$ and marginalize them out. The probability of a triple becomes:
\begin{align*}
p(T_{spo} = \text{true}) &= \int_{x_s, f_p, x_o \in S_{spo}} p(X_s = x_s) p(F_p = f_p) p(X_s = x_s) \,d x_s df_p dx_o\\
\text{with} \,\,S_{spo} &= \{(x_s, f_p, x_o) \mid f_p(x_s) = x_o\}
\end{align*}

\begin{align*}
p(\{X_i\}, \{F_p\}\mid T) = \frac{p(T\mid \{X_i\}, \{F_p\}) \prod_i p(X_i)\prod_p p(F_p)}{P(T)}
\end{align*}

@@

It is worth noting that we started from an impossible definition of the structure of our graph. The presumed unobserved true embeddings cannot exist for any graph contain multiple children or parents. Moreover, the system may be overdetermined: even with unique relations for every triple, there may still be no way to satisfy the whole knowledge graph in a single set of embeddings. Nevertheless, the Bayesian approach allows us to satisfy the resulting contradictions, by not choosing a single truth, but but providing a probability distribution over configurations that are each close to a correct one. In other words, the uncertainty about how the contradictions should be resolved is simply modeled as another probability, and folded into the model, combined with our other uncertainties.
